---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

```



## Loading and preprocessing the data
```{r}
library(knitr)
library(plyr)
library(dplyr)
library(ggplot2)

activity<-read.csv("Peer assessment 1/activity.csv",header = T)
head(activity)
```



## What is mean total number of steps taken per day?
```{r}
totalstepsperday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
head(totalstepsperday)
```
### Histogram of the total number of steps taken each day
```{r}
hist(totalstepsperday$steps, 
    main="Total Steps per Day", 
    xlab="Number of Steps per Day", 
    ylab = "Interval",
    col="cyan",
    breaks=50)
box()
```

### Mean of Total steps per day

```{r}
meansteps<-mean(totalstepsperday$steps)
meansteps
```
### Median of Total steps per day
```{r}
mediansteps<-median(totalstepsperday$steps)
mediansteps
```


## What is the average daily activity pattern?
### Five minute mean using steps to interval - FUN = mean instead of sum
```{r}
fivemin <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
plot(x = fivemin$interval, 
    y = fivemin$steps, 
    type = "l", 
    col = "red",
    xlab = "5-minute Intervals",
    ylab = "Average Steps Taken ~ Days",
    main = "Average Daily Activity Pattern")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxsteps <- fivemin$interval[which.max(fivemin$steps)]
maxsteps

```




## Imputing missing values
```{r}
sum(is.na(activity)) 
```
### Removing na values 
```{r}
newactivity <- activity
nas <- is.na(newactivity$steps)
avg_interval <- tapply(newactivity$steps, newactivity$interval, mean, na.rm=TRUE, simplify = TRUE)
newactivity$steps[nas] <- avg_interval[as.character(newactivity$interval[nas])]
names(newactivity)
```
### Histogram

```{r}
par(mfrow=c(1,2))
totalstepsperday2 <- aggregate(steps ~ date, data = newactivity, FUN = sum, na.rm = TRUE)
head(totalstepsperday2)
hist(totalstepsperday2$steps, 
    main = "Total Steps per Day (no-NA)", 
    xlab = "Number of Steps per Day", 
    ylab = "Interval",
    col="green",
    breaks=50)
##Histogram with the orginal dataset
hist(totalstepsperday$steps, 
    main="Total Steps per Day (Original)", 
    xlab="Number of Steps per Day", 
    ylab = "Interval",
    col="orange",
    breaks=50)
summary(totalstepsperday)
summary(totalstepsperday2)

```
## Are there differences in activity patterns between weekdays and weekends?

### Add the new weekend/weekday field

```{r}
activity2<- newactivity%>%
        mutate(typeofday= ifelse(weekdays(newactivity$date)=="Saturday" | weekdays(newactivity$date)=="Sunday", "Weekend", "Weekday"))
head(activity2)
fivemin2<- aggregate(steps ~ interval, data = activity2, FUN = mean, na.rm = TRUE)
head(fivemin2)

ggplot(activity2, aes(x =interval , y=steps, color=typeofday)) +
       geom_line() +
       labs(title = "Ave Daily Steps (type of day)", x = "Interval", y = "Total Number of Steps") +
       facet_wrap(~ typeofday, ncol = 1, nrow=2)
```

